 First of all, I would like to thank my advisor Daniel Whiteson and Andy Lankford for the guidance, support, trust, and freedom they have given me over the years. I am especially grateful for Daniel's continuous encouragement towards my growth into an independent researcher. 

I would also like to thank my postdocs Dan Guest and Mike Fenton. Thank you for all the big and small advice on my work and personal encouragements over the years- sometime in the form of French Kabob runs or pandemic zoom coffees sessions. A special thank you is also dedicated to Kate Pachal, for your generous support through my first major analysis, for teaching me so much that I know now, and for the afternoon cake time at R1. I will strive to be like what you all were to me to my future students.

I am grateful for my colleagues at CERN, especially Verena Martinez, Bill Murray, Ben Nachman, Bogdan Malaescu, Caterina Doglioni, Christopher Haynes, Ljiljana Morvaj, Heather Russell, and Kyle Cranmer. The end product of this thesis would not have the same rigor without your technical comments and contributions.

A gratitude is also extended to my informal advisors in Irvine, Tim Tait, Muchun Chen, Simona Murgia, and Michael Ratz for the intellectual support and advice. I also want to thank Flip Tanedo for all the coffee and restaurant runs in my first two years of grad school. I was nourished in the mind, in my stomach, and also in my caffeine level. I also want to thank Chase and Seyda for all the thought-provoking chats.

If I am allowed the indulgence to take it back to where the journey begins, I want to thank my middle science teacher Miss Chau, thank you for always having so much trust in my abilities and for showing me how practical and relatable science is. I also want to thank my math teacher Miss Fung: for both the lessons in math and outside of math. Some words outside of math took me many years more years to understand after class was over. I am grateful to my undergraduate friends and
``Senpais" David and Jason, for showing me the fun of solving technical problems with science and engineering and for being great friends. I am also thankful for research time from MC Chu in his lab. I learned not just the rigor of scientific research but also responsibility as a scientist to society. I aspire to continue my efforts of scientific outreach as a postdoc like you have done for so many. 

%NAthan , Chase, seyda matt and jennifer, 
It only took 3 continents to finish my Ph.D. These would not have been possible without the support of my friends across all three continents. Special shout out to Andy, Clare, Diane, Haidee, Alexis, Ros, Yanyi, CY, Rhoda, Alice, Joakim, Lily, Arianna, Freida, Kit, Ho, Amber, Tamara, Kiwi, Shelly, Danny, friends in Kindle Hills. My life in the last few years both instead and outside of physics really wouldn't have been the same without you.

Finally to my family: mom and dad, I hope I made you proud. To my brother Rene, I am lucky to have your support and to have you as a partner in crime through all of life's adventures.

This thesis is for all of you.

